The Dynamics of Anthropomorphic Effects in Human-Robot Interaction

### 2014-03-27 paper.tex
new bib item new paragraph our model section

\newcommand{\eg}{{\textit{e.g.~}}}
\newcommand{\etal}{{\textit{et al.~}}}
\newcommand{\ie}{{\textit{i.e.~}}}
\newcommand{\etc}{{\textit{etc.~}}}

@@

\begin{figure}[h!]
        \centering
        \begin{subfigure}[t]{0.48\columnwidth}
                \centering
                \includegraphics[width=0.6\textwidth]{ranger}
                \caption{The \emph{Ranger} Dominos experiment}
                \label{fig:ranger-expe}
        \end{subfigure}%
        ~ %add desired spacing between images, e. g. ~, \quad, \qquad etc.
          %(or a blank line to force the subfigure onto a new line)
        \begin{subfigure}[t]{0.48\columnwidth}
            \includegraphics[width=0.9\textwidth]{ranger-interactions-vs-qualitative-score}
            \caption{Qualitative level of anthropomorphic projections versus
            actual interactions with the robot}
            \label{fig:qualitative-score}
        \end{subfigure}
        \caption{Level of interaction with a semi-anthropomorphic robot is
        \emph{not} a predictor of anthropomorphic projections.}
    \label{fig:ranger}
\end{figure}

@@

The direction of causality in the artifact-centered view is that the non-human artifact 
(or a certain feature of it) encourages the human to anthropomorphize the artifact / 
agent. This perspective assumes that humans directly respond to life-like and social 
cues that the non-human agent or system emits, without thoughtful mental processing, 
by simply applying stereotypes and heuristics to it, and in turn apply 
human-human social schemas and norms to the occurring interactions.
This view is acceptable since humans naturally react to specific 
cues. From early childhood 
on, humans are inherently well-trained to perceive life \citep{epley_seeing_2007}, 
or to respond to social cues, such as facial expressions, social gestures or turn-taking. 
One has to keep in mind that these cues are of various senses and modalities in human-
human interaction, and consequently human-like design also addresses several senses. 
One of the most outstanding aspects is probably the outer appearance of an agent or 
artifact (in particular when it is unknown or unfamiliar). \cite{schmitz_concepts_2011} 
argues that an agent's outer appearance is likely to have an important impact on how 
the object is overall perceived. Consequently, if a robot's outer form resembles a 
human, the robot is likely to be treated similar to a human (\eg eye-contact is 
established or the robot is greeted).

@@

@@

The \emph{familiarization phase} starts at the peak of the \emph{novelty
effect}: anthropomorphic effects are at their maximum, the user thinks (s)he is
potentially facing an agent aiming at ``human-level intelligence'' (to take
McCarthy's words). This is a transient state that quickly vanishes, and the
projected anthropomorphism then starts to decrease while the human observes and
recognizes that the behaviour of the robot is generally \emph{predictable} and
possibly \emph{deceptive} (for instance, it talks but does not understand when
we talk ; it has eyes but does not recognize everyday objects we show ; etc.).

@@

As illustrated on Figure~\ref{fig:dynamics}, our model represents
\emph{disruptive behaviors} as either local increases or local decreases of
anthropomorphic effects in the familiarization phase (and at a lesser extend, in
the stabilization phase): because such behaviors are unexpected, a human
observer may interpret them either as the result of a richer deliberative
process, which in turn leads to the supposition of complex cognitive skills, or
conversely as a failure, reminding the human that the robot is ``just a
machine''. Section~\ref{sec:disruptive} discusses more in depth the impact of
the user's perception of disruptive behaviours, and their cognitive correlates.

@@

We proposed that \ICA is built on three factors: user's \emph{personality},
robot's \emph{design} and interaction \emph{purpose} (or \emph{interaction
context}). The user's personality and the context of use do also influence \SLA.
In particular, it appears that the user's level of acquaintance with
technologies plays an important role in long-term tendency to
anthropomophize~\cite{fink_living_2013} (people more familiar with technology
understand, and hence predict, better the behavior of the robot, which in turn
leads them more frequently to ultimately consider the robot as a simple tool).

@@

%From the human point of view, what are the brain mechanisms that could mediate anthropomorphism towards robots ?
\subsection{Neuroscience review of anthropomorphic brain mechanisms in human-robot interaction}
The neuroscientific study of human-robot interaction is still in its infancy. So far most studies have used robots as a way to investigate humans behaviour and brain responses to perception of an artefactual agent actions and emotional states but have not really tapped in the study of how a human behaviours and brain processes are influenced by its direct interaction with a robot (add ref, ref). 
We briefly review below the neuroscientific literature relevant to understand the neural basis of anthropomorphism.

%From an early age, humans develop a tendency to explain one's own and others'actions in terms of beliefs, desires and goals called theory of mind or mentalizing allowing the implicit attribution of intentions and other mental states to others \citep{Frith2003}.
%A system of brain areas have been linked to mentalizing that encompass an anterior region of the medial prefrontal cortex (MPFC), that may be the basis to distinguish between mental and physical states representation, the temporal lobes, that may allow access to social knowledge  and the posterior superior temporal sulctuc (STS), that might be involved in detection of agency \citep{Frith2003}.
Resonnance behaviors \citep{Rizzolatti1999} are mechanisms by which the brain areas involved in the internal representations of an action, an emotion or a sensation are equally recruited during the perception of another individual performing the same action or experiencing the same emotion or sensation.

\paragraph{Anthropomorphic response to motor behavior \\}
Resonnance behaviors have first been shown in the motor mirror neurons system, that is a neural network in motor areas that fire both during the execution of specific goal-oriented action and during the viewing of action directed toward the same goal~\cite{Rizzolatti1996, Kilner2009}.

There are good evidence showing that the human brain responds to both human and robotic motion by activating motor mirror neurons \citet{Gazzola2007, oberman_eeg_2007}, stressing that the perceiving the goal of an action is more important than the human-likeness of its kinematic to activate resonnance mechanisms.
%For example, fMRI study \citet{Gazzola2007} have found no significant difference in the activation of the mirror neurons system during the observation of an industrial robot or of a human performing the same action, suggesting that the goal of the action is more important than its kinematics to activate resonnance mechanisms. %mirror neurons

@@

\paragraph{Anthropomorphism and social behaviour}
In studying emotional perception in a human or a robot face using functional magnetic resonnance imagery (fMRI) \citet{Chaminade2010} found reduced activity in response to the robot stimuli than to the human stimuli in emotional brain areas known to have mirror properties. However if the subject was asked to specifically attend to the emotional content of the stimuli, processing the robot facial expresion, but not the human's, elicited a significant increase in the anterior part of the anterior frontal gyrus, a brain area considered as a marker of motor resonnance.

Howver, using electroencephalograpy another study have found that perception of emotions in a non-humanoid robot face affects the early brain processes of emotional expressions and the behavioral response of the participants the same way as did a human face \citep{Dubal2010}.

@@

Thus it seems that the display of resonance mechanisms for social behaviour in a human-robot interaction is less straightforward than for motor behaviour and may dependent more on the internal attitude and cognitive conceptions the human has of the robot. 
%Brain areas involved when attributing mental states to other human seem to be involved also in anthropomorphism . Indeed, individual scores of anthropomorphisms have been found to correlate with the grey matter volume of the left temporoparietal junction, a brain area activared during mentalizing \citet{cullen2013individual}

@@

The so-called \emph{level 1} is the instinctive, pre-cognitive identification of
living peers. That humans tend to anthropomorphize robots intuitively in this
pre-cognitive way
is supported by studies done by Rosenthal-von der Pütten
\textit{et al.}~\cite{rosenthal-vonderputten_experimental_2013} who investigated
the neural correlates of emotional reactions of humans towards a robot. {\it
Empathy} is characteristic of this stage~\cite{rosenthalvonderPutten2013neural}.
Anthropomorphism at this pre-cognitive stage might also be mediated by the motor and emotional resonnance mechanisms at the neuronal level.

After a longer observation period (typically including complete action sequences
of the robot) or short interaction (touching, short talk like greetings), we
suggest the human reaches the cognitive \emph{level 2}: in this phase, the human
starts building a behavioral and cognitive model of the robot that would support
both the observed and imagined capabilities of the robot.  The \emph{familiarity
thesis}~\cite{hegel_understanding_2008} supports the idea that the human
first projects onto the robot mental models of similar agents he/she is already
familiar with (ranging from animals to human adults, to pets and children). We 
hypothesize that the nature of the projected mental
model, as well as how deep the human engages in this projection, might be
driven by the same parameters as we mentioned for the \emph{initial capital of
anthropomorphism} (section~\ref{sec:ica}). At this stage the huma might also change his attitude towards the robot, paying more attention to social cue than to low-level behaviour which may reinforce the resonnance mechanisms.

@@

We present in this section the methodology and techniques we have relied on to
assess anthropomorphic projections in two unrelated studies, the Roomba
experiment~\cite{fink_living_2013}, and the Dominos experiment.

@@

\paragraph{Qualitative assessment of anthropomorphism} The main tools for
qualitative measurement of anthropomorphic effect are questionnaires. We have
explored both Likert-scale and open questions. In the Dominos experiment,
questionnaires were analyzed such as points are attributed for key answers,
leading to a qualitative score:

>>>>

\newcommand{\eg}{{\textit{e.g.~}}}
\newcommand{\etal}{{\textit{et al.~}}}
\newcommand{\ie}{{\textit{i.e.~}}}
\newcommand{\etc}{{\textit{etc.~}}}
\newcommand{\vs}{{\textit{vs.~}}}

@@

\begin{figure}[h!]
        \centering
        \begin{subfigure}[t]{0.48\columnwidth}
                \centering
                \includegraphics[width=0.6\textwidth]{ranger}
                \caption{The \emph{Ranger} Domino study}
                \label{fig:ranger-expe}
        \end{subfigure}%
        ~ %add desired spacing between images, e. g. ~, \quad, \qquad etc.
          %(or a blank line to force the subfigure onto a new line)
        \begin{subfigure}[t]{0.48\columnwidth}
            \includegraphics[width=0.9\textwidth]{ranger-interactions-vs-qualitative-score}
            \caption{Qualitative level of anthropomorphic projections versus
            actual interactions with the robot}
            \label{fig:qualitative-score}
        \end{subfigure}
        \caption{Level of interaction with a semi-anthropomorphic robot is
        \emph{not} a predictor of anthropomorphic projections.}
    \label{fig:ranger}
\end{figure}

@@

The direction of causality in the artifact-centered view is that the non-human artifact 
(or a certain feature of it) encourages the human to anthropomorphize the artifact / 
agent. This perspective assumes that humans directly respond to life-like and social 
cues that the non-human agent or system emits, without thoughtful mental processing, 
by simply applying stereotypes and heuristics to it, and in turn apply 
human-human social schemas and norms to the occurring interactions.
This view is acceptable since humans naturally react to specific 
cues. From early childhood 
on, humans are inherently well-trained to perceive life \citep{epley_seeing_2007}, 
or to respond to social cues, such as facial expressions, social gestures or turn-taking. Also \cite{takayama_perspectives_2012} states that humans have a propensity toward particular cues that seem to trigger automatic interpersonal social responses. 
One has to keep in mind that these cues are of various senses and modalities in human-
human interaction, and consequently human-like design also addresses several senses. 
One of the most outstanding aspects is probably the outer appearance of an agent or 
artifact (in particular when it is unknown or unfamiliar). \cite{schmitz_concepts_2011} 
argues that an agent's outer appearance is likely to have an important impact on how 
the object is overall perceived. Consequently, if a robot's outer form resembles a 
human, the robot is likely to be treated similar to a human (\eg eye-contact is 
established or the robot is greeted).

@@

\hl{Julia: I need to work on this next paragraph but I would like to keep it here as a note}

\textbf{In-the-moment vs. reflective perspectives on agency}\\
Perception of agency can be treated as anthropomorphism.
\citep{takayama_perspectives_2012} proposes to differentiate between a \textit{in-the-moment} and \textit{reflective} perspective to better understand how human users perceive agency in robots.\footnote{``The other important distinction between the two perspectives is the \hl{cognitive processes} that they engage.''} According to Takayama, an \textit{in-the-moment} perspective refers to one's most immediate (sometimes visceral) sense in a situation. In contrast, a \textit{reflective} perspective refers to one's sense of a situation upon more distanced cogitation and consideration. For instance, we reflectively might not perceive agency in a social robot, but it can feel quite differently in the moment of interaction. A critical dimension of agency is the perspective from which something seems to have agency. \cite{takayama_perspectives_2012} argues that neglecting to separate reflective perspectives from in-the-moment perspectives of agency is one of the major sources of confusion when people talk and write about anthropomorphism. There seems to be a disconnection between what people consciously perceive and how they respond to stimuli that they may not consciously perceive.
That means, in the initial phase of interacting with a novel device \textbf{(initialization)} people might not consciously respond but rather mindlessly \citep{nass_machines_2000}. Only after some time of \textbf{familiarization}, they might respond in a more reflective manner. This can be illustrated by the fact that people tend to deny interacting with computational systems as if they were people and yet they respond to computers in many ways that are remarkably similar to how they respond to people \citep{reeves_media_1996}.
\cite{takayama_perspectives_2012} also applies a cognitive viewpoint on the two different perspectives of in-the-moment \vs reflective to illustrate differences. She states that in-the-moment perceptions of agency are largely shaped by bottom-up perceptual processes, evoking very immediate responses (\eg to the cues emitted by the artifact's design, see Section \ref{sec:anthropomorphic-design}).
In contrast, according to Takayama, reflective perceptions are more often shaped by top-down processes because of the nature of reflective thought.

@@

The \emph{familiarization phase} starts at the peak of the \emph{novelty
effect}: anthropomorphic effects are at their maximum, the user thinks (s)he is
potentially facing an agent aiming at ``human-level intelligence'' (to take
McCarthy's words). This is a transient state that quickly vanishes, and the
projected anthropomorphism then starts to decrease while the human observes and
recognizes that the behavior of the robot is generally \emph{predictable} and
possibly \emph{deceptive} (for instance, it talks but does not understand when
we talk ; it has eyes but does not recognize everyday objects we show ; etc.).

@@

As illustrated on Figure~\ref{fig:dynamics}, our model represents
\emph{disruptive behaviors} as either local increases or local decreases of
anthropomorphic effects in the familiarization phase (and at a lesser extend, in
the stabilization phase): because such behaviors are unexpected, a human
observer may interpret them either as the result of a richer deliberative
process, which in turn leads to the supposition of complex cognitive skills, or
conversely as a failure, reminding the human that the robot is ``just a
machine''. Section~\ref{sec:disruptive} discusses more in depth the impact of
the user's perception of disruptive behaviors, and their cognitive correlates.

@@

We proposed that \ICA is built on three factors: user's \emph{personality},
robot's \emph{design} and interaction \emph{purpose} (or \emph{interaction
context}). The user's personality and the context of use do also influence \SLA.
In particular, it appears that the user's level of acquaintance with
technologies plays an important role in long-term tendency to
anthropomorphize~\cite{fink_living_2013} (people more familiar with technology
understand, and hence predict, better the behavior of the robot, which in turn
leads them more frequently to ultimately consider the robot as a simple tool).

@@

%From the human point of view, what are the brain mechanisms that could mediate anthropomorphism towards robots ?
\subsection{Neuroscience review of anthropomorphic brain mechanisms in human-robot interaction}
The neuroscientific study of human-robot interaction is still in its infancy. So far most studies have used robots as a way to investigate humans behavior and brain responses to perception of an artefactual agent actions and emotional states but have not really tapped in the study of how a human behaviors and brain processes are influenced by its direct interaction with a robot (add ref, ref). 
We briefly review below the neuroscientific literature relevant to understand the neural basis of anthropomorphism.

%From an early age, humans develop a tendency to explain one's own and others'actions in terms of beliefs, desires and goals called theory of mind or mentalizing allowing the implicit attribution of intentions and other mental states to others \citep{Frith2003}.
%A system of brain areas have been linked to mentalizing that encompass an anterior region of the medial prefrontal cortex (MPFC), that may be the basis to distinguish between mental and physical states representation, the temporal lobes, that may allow access to social knowledge  and the posterior superior temporal sulctuc (STS), that might be involved in detection of agency \citep{Frith2003}.
Resonance behaviors \citep{Rizzolatti1999} are mechanisms by which the brain areas involved in the internal representations of an action, an emotion or a sensation are equally recruited during the perception of another individual performing the same action or experiencing the same emotion or sensation.

\paragraph{Anthropomorphic response to motor behavior \\}
Resonance behaviors have first been shown in the motor mirror neurons system, that is a neural network in motor areas that fire both during the execution of specific goal-oriented action and during the viewing of action directed toward the same goal~\cite{Rizzolatti1996, Kilner2009}.

There are good evidence showing that the human brain responds to both human and robotic motion by activating motor mirror neurons \citet{Gazzola2007, oberman_eeg_2007}, stressing that the perceiving the goal of an action is more important than the human-likeness of its kinematic to activate resonance mechanisms.
%For example, fMRI study \citet{Gazzola2007} have found no significant difference in the activation of the mirror neurons system during the observation of an industrial robot or of a human performing the same action, suggesting that the goal of the action is more important than its kinematics to activate resonnance mechanisms. %mirror neurons

@@

\paragraph{Anthropomorphism and social behavior}
In studying emotional perception in a human or a robot face using functional magnetic resonance imagery (fMRI) \citet{Chaminade2010} found reduced activity in response to the robot stimuli than to the human stimuli in emotional brain areas known to have mirror properties. However if the subject was asked to specifically attend to the emotional content of the stimuli, processing the robot facial expression, but not the human's, elicited a significant increase in the anterior part of the anterior frontal gyrus, a brain area considered as a marker of motor resonance.

However, using electroencephalography another study have found that perception of emotions in a non-humanoid robot face affects the early brain processes of emotional expressions and the behavioral response of the participants the same way as did a human face \citep{Dubal2010}.

@@

Thus it seems that the display of resonance mechanisms for social behavior in a human-robot interaction is less straightforward than for motor behavior and may dependent more on the internal attitude and cognitive conceptions the human has of the robot. 
%Brain areas involved when attributing mental states to other human seem to be involved also in anthropomorphism . Indeed, individual scores of anthropomorphisms have been found to correlate with the grey matter volume of the left temporoparietal junction, a brain area activared during mentalizing \citet{cullen2013individual}

@@

The so-called \emph{level 1} is the instinctive, pre-cognitive identification of
living peers. That humans tend to anthropomorphize robots intuitively in this
pre-cognitive way
is supported by studies done by Rosenthal-von der Pütten
\textit{et al.}~\cite{rosenthal-vonderputten_experimental_2013} who investigated
the neural correlates of emotional reactions of humans towards a robot. {\it
Empathy} is characteristic of this stage~\cite{rosenthalvonderPutten2013neural}.
Anthropomorphism at this pre-cognitive stage might also be mediated by the motor and emotional resonance mechanisms at the neuronal level.

After a longer observation period (typically including complete action sequences
of the robot) or short interaction (touching, short talk like greetings), we
suggest the human reaches the cognitive \emph{level 2}: in this phase, the human
starts building a behavioral and cognitive model of the robot that would support
both the observed and imagined capabilities of the robot.  The \emph{familiarity
thesis}~\cite{hegel_understanding_2008} supports the idea that the human
first projects onto the robot mental models of similar agents he/she is already
familiar with (ranging from animals to human adults, to pets and children). We 
hypothesize that the nature of the projected mental
model, as well as how deep the human engages in this projection, might be
driven by the same parameters as we mentioned for the \emph{initial capital of
anthropomorphism} (section~\ref{sec:ica}). At this stage the human might also change his attitude towards the robot, paying more attention to social cue than to low-level behavior which may reinforce the resonance mechanisms.

@@

We present in this section the methodology and techniques we have relied on to
assess anthropomorphic projections in two unrelated studies, the Roomba
experiment~\cite{fink_living_2013}, and the Domino study.

@@

\paragraph{Qualitative assessment of anthropomorphism} The main tools for
qualitative measurement of anthropomorphic effect are questionnaires. We have
explored both Likert-scale and open questions. In the Domino study,
questionnaires were analyzed such as points are attributed for key answers,
leading to a qualitative score:
